---
title: "Introduction to ggpic"
author: "Victor Granda"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to ggpic}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

`ggpic` extends `ggplot2` package adding geoms based on the the default ones, but
using raster objects instead of rect elements. This way you can make a barplot
with trees, buildings or leaves instead bars.

## Intro

disclaimer about why ggpic, no using it in serious business and its potential
in presentations and posters.

## Installation

`ggpic` is only available from GitHub at the moment. You can install it using
`devtools` package:

```{r, eval=FALSE}
# install devtools if necessary
install.packages('devtools')

# install ggpic
devtools::install_github('MalditoBarbudo/ggpic', build_vignettes = TRUE)
```

## Usage

`geom_bar_pic` works as `bar_geom`, so it needs at least an `x` aesthetic. It
also uses a `pic` aesthetic to indicate the image to use.

```{r basic_plot, message = FALSE, fig.align='center', fig.width=6}
# libraries
library(ggplot2)
library(ggpic)
library(dplyr)

# pic
leaf <- as.matrix(as.raster(rsvg::rsvg(
  system.file('pics', 'leaf.svg', package = 'ggpic')
)))

# plot
iris %>%
  ggplot(aes(x = Species)) +
  geom_bar_pic(pic = 'leaf')
```

***

It also accepts an `y` aesthetic, in combination with `stat = identity`, to map
y value to a variable value:

```{r y_aes, message = FALSE, fig.align='center', fig.width=6}
# plot
iris %>%
  group_by(Species) %>%
  summarise_all(funs(mean)) %>%
  ggplot(aes(x = Species)) +
  geom_bar_pic(aes(y = Petal.Width), stat = 'identity', pic = 'leaf')
```

***

`geom_bar_pic` can use the `fill` and the `alpha` aesthetics, and also the
`width` parameter from `geom_bar`. This could be useful to avoid excesive
pic deformation:

```{r fill_alpha_width, message = FALSE, fig.align='center', fig.width=6}
# width
iris %>%
  group_by(Species) %>%
  summarise_all(funs(mean)) %>%
  ggplot(aes(x = Species)) +
  geom_bar_pic(aes(y = Petal.Width),
               stat = 'identity', pic = 'leaf', width = 0.6)

# fill
iris %>%
  group_by(Species) %>%
  summarise_all(funs(mean)) %>%
  ggplot(aes(x = Species, fill = Species)) +
  geom_bar_pic(aes(y = Petal.Width),
               stat = 'identity', pic = 'leaf', width = 0.6)

# alpha
iris %>%
  group_by(Species) %>%
  summarise_all(funs(mean)) %>%
  ggplot(aes(x = Species, fill = Species)) +
  geom_bar_pic(aes(y = Petal.Width),
               stat = 'identity', pic = 'leaf', width = 1.3, alpha = 0.5)
```

As it can be seen in the last two plots, `geom_bar_pic` also provides a
legend with the pic and the corresponding fill colour (see `?draw_key_pic`).

***

The `pic` aesthetic, as any other, can be mapped to a variable. In this case it is
necessary to provide a scale for the `pic` aesthetic with the `scale_pic_manual`
function:

```{r scale, message = FALSE, fig.align='center', fig.width=6}
# data
data <- data.frame(
  Site = rep(LETTERS[1:3], 2),
  Species = rep(c('Pine', 'Ash'), each = 3),
  Height = rpois(6, 15),
  Deep = rpois(6, 2.2),
  stringsAsFactors = FALSE
)

# pics
pine <- as.matrix(as.raster(rsvg::rsvg(
  system.file('pics', 'pine_tree_2.svg', package = 'ggpic')
)))
ash <- as.matrix(as.raster(rsvg::rsvg(
  system.file('pics', 'deciduous_tree_2.svg', package = 'ggpic')
)))

# plot
ggplot(data,
       aes(x = Site, pic = Species, fill = Species)) +
  geom_bar_pic(aes(y = Height), stat = 'identity') +
  scale_pic_manual(values = c('ash', 'pine'))
```

## Pics

Some info about the raster objects